Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-04 data-independence-EPQ.tex
add tex

\section{Data-Independence of Priority Queue}
\label{sec:data-independence of priority queue}

Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain. In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for priority queues.

Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of priority queue is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$. The definition extends to (sets of) executions and histories. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

@@

A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
The set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

@@

The following lemma states that, when checking that a data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification, it is enough to do so for data-differentiated executions, similar as that in \cite{Abdulla:2013}, where the notion of data-independence and differentiated in \cite{Wolper:1986} is used. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of PQ}.

>>>>

\section{Data-Independence of Extended Priority Queue}
\label{sec:data-independence of extended priority queue}

Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain. In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for extended priority queues.

Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of extended priority queue is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$. The definition extends to (sets of) executions and histories. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

@@

A renaming function $r$ for extended priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged. The intuitive explanation is that renaming items will not influence the executions of program, while renaming priorities may influence the executions of program.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
A set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

@@

The following lemma states that, when checking that a data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification, it is enough to do so for data-differentiated executions, similar as that in \cite{Abdulla:2013}, where the notion of data-independence and differentiated in \cite{Wolper:1986} is used. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.

### 2017-04-04 preliminaries-EPQ.tex
add tex

\redt{We fix several (possibly infinite) set $\mathbb{D}_1,\ldots$ of data values}, and a finite set $\mathbb{M}$ of methods. %We consider that methods have exactly one argument, or one return value.
\redt{Return values are transformed into argument values for uniformity \footnote{\redt{Method return values are guessed nondeterministically, and validated at return points. This can be handled using the assume statements of typical formal specification languages, which only admit executions satisfying a given predicate. The argument value for methods without argument or return values, or with fixed argument/return values, is ignored.}}.}
We identify a subset $\mathbb{M}_{\textit{in}} \subseteq \mathbb{M}$ of input methods in order to differentiated methods taking an argument (e.g., the $\textit{put}$ method which inserts a argument value into a priority queue) from the other methods (e.g., the $\textit{rm}$ method which doesn't take an argument, and returns the item with maximal priority of a queue). A method-event is composed of a method $m \in \mathbb{M}$ and \redt{several data value $x_1 \in \mathbb{D}_1,\ldots$, and is denoted $m(x_1,\ldots)$}. We define the concatenation of method-event sequences $u \cdot v$ in the usual way, and $\epsilon$ denotes the empty sequence.

>>>>

\redt{We fix several (possibly infinite) set $\mathbb{D}_1,\ldots$ of data values}, and a finite set $\mathbb{M}$ of methods. %We consider that methods have exactly one argument, or one return value.
%\redt{Return values are transformed into argument values for uniformity \footnote{\redt{Method return values are guessed nondeterministically, and validated at return points. This can be handled using the assume statements of typical formal specification languages, which only admit executions satisfying a given predicate. The argument value for methods without argument or return values, or with fixed argument/return values, is ignored.}}.}
We identify a subset $\mathbb{M}_{\textit{in}} \subseteq \mathbb{M}$ of input methods in order to differentiated methods taking an argument (e.g., the $\textit{put}$ method which inserts a argument value into a priority queue) from the other methods (e.g., the $\textit{rm}$ method which doesn't take an argument, and returns the item with maximal priority of a queue). A method-event is composed of a method $m \in \mathbb{M}$ and \redt{several data value $x_1 \in \mathbb{D}_1,\ldots$, and is denoted $m(x_1,\ldots)$}. We define the concatenation of method-event sequences $u \cdot v$ in the usual way, and $\epsilon$ denotes the empty sequence.
